Spatial evolution of solar wind discontinuities in the outer heliosphere

### 2024-11-22 article.qmd
doc: Add waiting time results

The solar wind plasma state, as evidenced by the sunspot number (referenced in panel (c) of @fig-overview), plays a crucial role in understanding the dynamics of discontinuities. During the initial phase of the JUNO mission, the sunspot number reached its peak, indicating a period of heightened solar activity. However, by the end of the mission's cruise phase, there was a significant decline in solar activity. This variation underscores the importance of calibrating the discontinuity properties in relation to solar activity levels to account for temporal fluctuations.

@@

More than ? IDs are identified in the solar wind data from Juno, with ? IDs identified in 5 AU. Then for each discontinuity identified, we calculate the distance matrix of the time series sequence (distance between each pair of magnetic field vectors) to determine the leading edge and trailing edge of the discontinuity. After that, we use the minimum or maximum variance analysis (MVA) analysis [@sonnerupMinimumMaximumVariance1998; @sonnerupMagnetopauseStructureAttitude1967] to transform the magnetic field into the boundary normal (LMN) coordinate system. The main (most varying) magnetic field component, $B_l$, is then fitted by a hyperbolic tangent functions to extract the parameters to properly describe the discontinuity @eq-fit. Then we combine the magnetic field data and plasma data to obtain the thickness and the current density of the discontinuity.

$$
B_l(t) = B_{l0} \tanh\left(\frac{t - t_0}{\Delta T}\right) + c_l
$${#eq-fit}

where $B_{l0}$ is the amplitude of the magnetic field change in the maximum variance direction, $t_0$ is the time of discontinuity, $\Delta T$ is the transition time, and $c_l$ is the offset of the magnetic field change (in the case of an ideal discontinuity with perfect rotational symmetry, $c_l$ should theoretically be zero. However, in the context of data processing, incorporating $c_l$ can enhance the fitting accuracy).

@@

@@

where $μ_0$ is the vacuum permeability and $V_n$ is the normal component of the proton flow velocity. For fitting method, the derivative of the magnetic field could be estimated as $\max(d B_l / d t) = B_{l0} / \Delta T$. This estimate is more robust than the direct calculation of the derivative of the magnetic field, because the fitting method is less sensitive to the noise in the data and the time resolution of the data (see discussion in the appendix).

@@

$$
d = V_n \Delta T
$$

where $\Delta T$ is the transition time derived from the fitting method.

@@

The transition time $\Delta T$ is determined in one second resolution data for all our missions to reduce the system error. (Note that highest time resolution of the data is not needed for the transition time determination by fitting method).

>>>>

The solar wind plasma state, as evidenced by the sunspot number (referenced in panel (c) of @fig-overview), plays a crucial role in understanding the dynamics of discontinuities. During the initial phase of the JUNO mission, the sunspot number peaked, indicating a period of heightened solar activity. However, by the end of the mission's cruise phase, there was a significant decline in solar activity. This variation underscores the importance of calibrating the discontinuity properties in relation to solar activity levels to account for temporal fluctuations.

@@

More than ? IDs are identified in the solar wind data from Juno, with ? IDs identified in 5 AU. Then for each discontinuity identified, we calculate the distance matrix of the time series sequence (distance between each pair of magnetic field vectors) to determine the leading edge and trailing edge of the discontinuity. After that, we use the minimum or maximum variance analysis (MVA) analysis [@sonnerupMinimumMaximumVariance1998; @sonnerupMagnetopauseStructureAttitude1967] to transform the magnetic field into the boundary normal (LMN) coordinate system. The main (most varying) magnetic field component, $B_l$, is then fitted by a hyperbolic tangent function to extract the parameters to properly describe the discontinuity @eq-fit. Then we combine the magnetic field data and plasma data to obtain the thickness and the current density of the discontinuity.

$$
B_l(t) = B_{l0} \tanh\left(\frac{t - t_0}{\Delta t}\right) + c_l
$${#eq-fit}

where $B_{l0}$ is the amplitude of the magnetic field change in the maximum variance direction, $t_0$ is the time of discontinuity, $\Delta t$ is the transition time, and $c_l$ is the offset of the magnetic field change (in the case of an ideal discontinuity with perfect rotational symmetry, $c_l$ should theoretically be zero. However, in the context of data processing, incorporating $c_l$ can enhance the fitting accuracy).

@@

## Waiting time

To comprehensively understand the radial evolution of the occurrence rate, particularly the average occurrence rate, it is essential to determine the statistical distributions underlying the observed samples. This enables the derivation of meaningful inferences. In this context, we derived the probability distribution of the waiting time $τ$ (also referred to as the interarrival interval, as defined in @tsurutaniInterplanetaryDiscontinuitiesTemporal1979) at various radial distances.

To model the data, we employ the Weibull distribution due to its flexibility in capturing varying occurrence rates. This choice allows for a more accurate representation and an improved fit compared to the exponential distribution previously used in @tsurutaniInterplanetaryDiscontinuitiesTemporal1979. The probability density function of the Weibull distribution is given as follows:

$$
f(τ; α, θ) = \frac{α}{θ} \left( \frac{τ}{θ} \right)^{α-1} e^{-(τ/θ)^α}, \quad τ \ge 0
$$

Here, $α$ and $θ$ represent the shape and scale parameters, respectively. The complementary cumulative distribution function of the Weibull distribution $F(τ; α, θ) = 1 - \mathrm{e}^{-(τ/θ)^α}$ is a stretched exponential function.

::: {#fig-waitingTime}
![](figures/waiting_time_distribution)

Waiting time distribution, along with the fitted Weibull distribution at different radial distances.
:::

The observed waiting time distribution, along with the fitted Weibull distribution, is shown in @fig-waitingTime. For $α < 1$, the likelihood of an event occurring decreases with increasing waiting time. This behavior aligns with our expectations for the discontinuity occurrence rate in response to the evolving state of the solar wind. Across the largest range of waiting times, the observed probabilities exceed those predicted by the fitted distribution. This discrepancy may be attributed to data gaps, which artificially extend the observed waiting times between discontinuities.

The mean waiting time $\bar{τ}$ can be determined by integrating the probability density function $\bar{τ} = \int_0^\infty τ f(τ) dτ$. The mean occurrence rate, in turn, is simply the reciprocal of the mean waiting time. Notably, the mean value derived from the fitted Weibull distribution is more reliable, as it mitigates the influence of unusually large waiting times caused by data gaps.

@@

where $μ_0$ is the vacuum permeability and $V_n$ is the normal component of the proton flow velocity. For fitting method, the derivative of the magnetic field could be estimated as $\max(d B_l / d t) = B_{l0} / \Delta t$. This estimate is more robust than the direct calculation of the derivative of the magnetic field, because the fitting method is less sensitive to the noise in the data and the time resolution of the data (see discussion in the appendix).

@@

$$
d = V_n \Delta t
$$

where $\Delta t$ is the transition time derived from the fitting method.

@@

The transition time $\Delta t$ is determined in one second resolution data for all our missions to reduce the system error. (Note that highest time resolution of the data is not needed for the transition time determination by fitting method).
